---
title: "Dania"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2024-05-13"
---

```{r}
# Load necessary libraries
library(ggplot2)

# Load the Orange dataset
data(Orange)

# Create a ggplot with smoothing models
ggplot(Orange, aes(x = age, y = circumference, color = Tree)) +
  geom_smooth(colour="green") +  # Add smoothing model
  labs(title = "Growth of Orange Trees",
       x = "Age (days)",
       y = "Circumference (mm)",
       color = "Tree") +
  theme_minimal()

```

## Detrending and Time-series decomposition:

```{r}

# Load the "AirPassengers" dataset
data("AirPassengers")

# Convert the dataset into a time-series object
passengers_ts <- ts(AirPassengers, frequency = 12)

# Plot the time series data to observe the trend and seasonal patterns
plot(passengers_ts, main = "International Airline Passengers Over Time",
     xlab = "Year", ylab = "Number of Passengers")
```

## Geospatial Data

### a) various projections

```{r}
library(maps)
library(mapdata)
library(ggplot2)

# Load world map data
world_map <- map_data("world")

# a) Various projections
# Plot using the default projection
ggplot() +
  geom_map(data = world_map, map = world_map,
           aes(x = long, y = lat, map_id = region),
           fill = "white", color = "black") +
  theme_void()

# Plot using a different projection (e.g., Robinson)
ggplot() +
  geom_map(data = world_map, map = world_map,
           aes(x = long, y = lat, map_id = region),
           fill = "white", color = "black") +
  coord_map("mercator") +
  theme_void()

```

### b) layers visualizing

```{r}
library(maps)
library(ggplot2)

# Load world map data with higher resolution
world_map <- map_data("world")

# Plot continents and country borders
ggplot() +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill = "lightblue", color = "black") +
  ggtitle("Continents and Country Borders")
```
### c) Choropleth mapping

```{r}
# Load map data
map_mercator <- map_data("world")

# Generate random data for choropleth mapping
set.seed(123)  # for reproducibility
random_data <- data.frame(region = unique(map_mercator$region), value = rnorm(length(unique(map_mercator$region))))

# Merge data with map data
merged_data <- merge(map_mercator, random_data, by = "region", all.x = TRUE)

# Plot choropleth map
ggplot() +
  geom_polygon(data = merged_data, aes(x = long, y = lat, group = group, fill = value), color = "black") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  ggtitle("Choropleth Map of Random Data")
```
### d) Cartograms

```{r}
library(sf)
library(maps)
library(ggplot2)

# Load world map data
world_map <- map("world", fill = TRUE, plot = FALSE)

# Convert to sf object
world_sf <- st_as_sf(world_map)

# Plot the world map with red outlines
ggplot() +
  geom_sf(data = world_sf, color = "blue",aes(fill = "pop")) +  # Use color = "red" for red outlines
  theme_void()
```

## Probability as Frequency:

```{r}
# Load the InsectSprays dataset
data("InsectSprays")

# Create a frequency table to count the number of insects killed by each insecticide
frequency_table <- table(InsectSprays$spray)

# Calculate the probability of each insecticide
probabilities <- prop.table(frequency_table)

# Plot the probability as frequency plot
barplot(probabilities, main = "Probability as Frequency Plot of Insecticides",
        xlab = "Insecticides", ylab = "Probability")
```
## Quantile Dot plot Representation:

```{r}
# Step 1: Load the required packages
library(ggplot2)

# Step 2: Load the mtcars dataset
data(mtcars)


# Step 3: Create a quantile dot plot approximating the shape of the density
ggplot(mtcars, aes(x = factor(cyl), y = mpg)) +
  geom_dotplot(binaxis = "y", stackdir = "center", fill = "blue", binwidth = 1) +
  stat_summary(fun = median, geom = "point", color = "red", size = 3) +
  stat_summary(fun = mean, geom = "point", color = "green", size = 3) +
  geom_dotplot(aes(y = ..density..), binaxis = "y", stackdir = "center", fill = "grey", alpha = 0.5, binwidth = 0.5) +
  labs(title = "Quantile Dot Plot approximating Density of MPG by Cylinder Count",
       x = "Cylinder Count",
       y = "MPG") +
  theme_minimal()
```
## Confidence intervals:

```{r}
# Load the iris dataset
data(iris)

# Compute the mean and confidence intervals for Sepal Length
mean_sepal_length <- mean(iris$Sepal.Length)
ci_80 <- t.test(iris$Sepal.Length, conf.level = 0.80)$conf.int
ci_95 <- t.test(iris$Sepal.Length, conf.level = 0.95)$conf.int
ci_99 <- t.test(iris$Sepal.Length, conf.level = 0.99)$conf.int

# Print the results
cat("Mean Sepal Length:", mean_sepal_length, "\n")
cat("80% Confidence Interval:", ci_80, "\n")
cat("95% Confidence Interval:", ci_95, "\n")
cat("99% Confidence Interval:", ci_99, "\n")
```
```{r}
weather_data <- read.csv("weather.csv", stringsAsFactors = FALSE)
head(weather_data)
```
## my dataset

A time series plot is chosen to visualize the precipitation data over time.
It helps in understanding the overall trend, seasonality, and any other patterns present in the data.


```{r}
# Load required libraries
library(ggplot2)
library(forecast)


# Convert Date.Full to Date format
weather_data$Date.Full <- as.Date(weather_data$Date.Full)

# Plot the time series of precipitation
ggplot(weather_data, aes(x = Date.Full, y = Data.Precipitation)) +
  geom_line() +
  labs(title = "International Airline Passengers Over Time",
       x = "Date", y = "Precipitation") +
  theme_minimal()

# Time-series decomposition
ts_data <- ts(weather_data$Data.Precipitation, frequency = 12)
decomposed <- decompose(ts_data)
plot(decomposed)
```

